A Global Market for Marine Conservation

### 2021-06-15 draft.Rmd
Updated intro and part of methods, added references

---
title: "A Global Market for Marine Conservation"
subtitle: "Draft"
# https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown
author:
  - Juan Carlos Villaseñor-Derbez:
      email: juancarlos@ucsb.edu
      institute: bren
      correspondence: true
  - name: Darcy Bradley
    institute: bren
  - name: Christopher Costello
    institute: bren
  - name: Juan Mayorga
    institute: bren
  - name: Andrew Plantinga
    institute: bren
institute:
  - bren: Bren School of Environmental Science and Management, University of California Santa Barbara
date: "Prepared on `r format(Sys.time(), '%B %d, %Y')`"
abstract: "Humanity depends on the ocean for livelihood, subsitance, and way of life. Yet, many of our activities continiously pose a threat to marine populations and ecosystems. The implementation of Marine Protected Areas (MPAs), where extractive activities are regulated or prohibited on the basis of spatial deliniations, are one of many ways to manage the oceans. Current intenrational agreements call to achieve a certain amount of within-country conservation. Here we explore the benefits from implementing a market for conservation in which nations trade conservation towards a global target. We find that such a market reduces the costs of conservation by 46%, relative to a BAU scenario without a market where each country protects a fixed portion of their waters. When creating 12 distinct markets based on biogeography regionalization, we find that costs are reduced by 34.8%."
fontsize: 12pt
output:
  bookdown::pdf_document2:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: no
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{pdfpages}
  - \usepackage{placeins}
---

@@

# Important sentences

Importantly, the quality of a patch can be defined independently of other pixels. This is true for habitat suitability, in which the abiotic characteristics that determine the suitability of a patch are independent of the protected status of any patch.

It is important that we acknowledge the equity concerns of any conservation intenrvention. Under the current approach to conservation, nations face two options: to conserve 30% in situ, or not engage in conservation at all. This market-based approach allows for a third option: it allows nations to decide how much to conserve in their waters.

Ocean grabbing, the term that relates to the loss of soveraignty in the ocean (see Benett) may be perceived as a problem. However, nations are not forced to conserve in situ, and will only do so if they are approrpiately compensated for it.

All ecosystems are euqlly important for the subsistance of nature and species that reside in them. But a coral reef is not comparable to a kelp forest.  A market design that doesn't acknowledge these differences may unintendendly bias conservation towards different ecosystems or geographic regions. When outcomes of a market are restricted to a geographic extent, such as air pollution markets, a buble policy is implemented. We borrow from this principle and use biogeographic regionalization principles to create ecologically-coherent market segments where nations trade similar habitat. For example, in the temperate eastern pacific, Mexico and the US will trade a same type of habitat, while in the Tropical Atlantic they wil.

Having tackled these three limitations (spatially-independent measure, x, and market bubbles), we proceed to simulate the trade outcomes of such a market. We evaluate the outcomes across a range of target coverages (1 - 100), and find that a market always results in reduced costs. The absolute and relative magnitude of the costs-savings changes with protection target, peaking at between X and Y, depending on of the choice market segmentation. Intuitively, increasing teh number of market segments reduces the opportunity for bargaining and reduces the gains from trade, but ensures the persistance of different habitats.

We pay especial attention to goals of interest, such as protecting 30% of the marine area (habitat suitability-weighted) within all national jurisdictions. We find that a market-based approach to conservation would reduce costs by X%, and that Y% less area would be needed.

# Introduction

Nations have committed to protect a fixed portion of their EEZs in order to reduce and control anthropogenic impacts in the marine environment. Aichi target 11^[Aichi target 11 states that by 2020 10% of coastal and marine areas should be conserved through protected areas.] and target 14.5 of the United Nations Sustainable Development Goals^[Similarly, target 14.5 of the United Nations Sustainable Development Goals is to conserve at least 10% of coastal and marine areas.] aim to protect 10% of coastal and marine areas. Some more ambitious goals call for 30%, such as that from the World Parks Congress^[the International Union for the Conservation of Nature (IUCN) has suggested that at least 30% of each marine habitat should be protected in areas with no extractive activities.].

These goals ultimately aim to protect a global amount of biodiversity by spatially abating risks. However, costs and benefits are heterogeneous in space, and even conservation-minded countries might struggle to meet these targets if costs are prohibitive. But what if a nation could pay to conserve elsewhere, where the costs are lower? This setup is similar to that of pollution abatement in a multi-firm setting. A large body of theoretical and empirical literature has shown that an efficient way to achieve a level of pollution abatement is by leveraging differences in firm's marginal cost curves and incentivizing trade of emission permits.

Just as any two firms are capable of reducing their emissions by a given amount, two nations are capable of implementing MPAs with the same biodiversity value. However one nation (or firm) may be able to do so at a lower cost. A market approach may be the most efficient (*i.e.* cost-effective) way of balancing these differences in costs and benefits, and ultimately allow us to reach the desired targets at a reduced cost.

Under such a market, what would be the gains from trade? For the same amount of conservation as in the BAU scenario, how much are the costs reduced? How does spatially dividing the market (say, by ecosystem or ocean) change the outcome? This are indeed the questions that I seek to answer.

# Methods

We combine data on biodiversity benefit and costs to generate supply curves for each nation, as well as a global biodiversity supply curve. While in practice this approach can be used for any given biodiversity target, I produce an example using the 30% area-based target (*i.e.* We translate this area-based target into biodiversity units). We then find the trading price in the global supply curve and use the country-level supply curves to calculate in-situ and ex-situ conservation for every country. We can then calculate the differences in costs, and estimate the gains from trade.

@@

The results presented so far assume that a unit of biodiversity is worth the same anywhere around the world. However, we might want to ensure that no single type of habitat or region is left out. To overcome this, markets can be established at different spatial scales. Here, we use different scales of zoning in the marine environment (*i.e.* Biogeographic zoning). \citet{spalding_2007} divided the oceans using a hierarchical approach with Realms (n = 12) containing Provinces (n = 60), which in turn contain Ecoregions (n = 219).

@@

![\label{fig:realm-map}Map of the 12 realms, as defined by \citet{spalding_2007}.](results/img/realm_map.png)

![\label{fig:eezs-per-realm} Number of nations contained within each of the realms described by \citet{spalding_2007}. The vertical dashed line lies at 2,  the minimum number of nations required for trade to occur.](results/img/eezs_per_realm.png)

@@

@@

We must first identify the costs and benefits of conservation. Conserving biodiversity may provide benefits to society. For example, areas with higher species richness and biomass are more attractive to tourism \citep{williams_2000}. Species richness also leads to increased stability of ecosystems \citep{tilman_1995}, and may reduce exposure to pathogens \citep{lamb_2017}. The quantification is not straightforward, but we can assume that biodiversity indices are positively correlated with their benefits they provide (although saturation may occur past a given point).

@@

On the other hand, we need to quantify the costs associated to each pixel. An often-cited cost of implementing MPAs is that of losses in fisheries revenues \citep{smith_2010}. In this case, we define costs as losses in fisheries landings ^[I intend to use losses in fisheries revenues in the future. This will involve species-level changes in landings, as well as species-level price data. Essentially, the global losses in landings used here will be weighed by species price data.]. Since protection of some areas may lead to rebuilding fisheries, parts of the ocean might have negative costs. This means that there are gains to be made in conserving them. A map of global costs of conservation is shown in Figure \ref{fig:landing-losses}. As before, these data do not consider the transient dynamics, and represent the equilibrium state of protecting any given pixel.

@@

Although there are important caveats, these data provide us with a way of calculating the costs and benefits of protecting any place under a nation's jurisdiction. While protecting the high seas may provide many benefits \citep{white_2014, sumaila_2015}, our market-based approach requires enforceable property rights, such as those offered by Exclusive Economic Zones. Moreover, biodiversity benefit is higher within Exclusive Economic Zones, where most of the species live \citep{mayorga_2020}. Therefore, we will ignore areas beyond national jurisdiction.

@@

<!-- MAYBE:: A meta-analysis of simulation studies suggests that targets ranging from biodiversity protection to food provisioning can be indeed achieved by protecting ~30% of our coastal oceans \citep{oleary_2016,cabral_2019}. However, these area-based targets ultimately imply a desired level of biodiversity benefit. -->

>>>>

---
title: "A Global Market for Marine Conservation"
subtitle: "Draft"
# https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown
author:
  - Juan Carlos Villaseñor-Derbez:
      email: juancarlos@ucsb.edu
      institute: bren
      correspondence: true
  - name: Darcy Bradley
    institute: bren
  - name: Christopher Costello
    institute: bren
  - name: Juan S. Mayorga-Henao
    institute: bren
  - name: Andrew Plantinga
    institute: bren
institute:
  - bren: Bren School of Environmental Science and Management, University of California Santa Barbara
date: "Prepared on `r format(Sys.time(), '%B %d, %Y')`"
abstract: "Humanity depends on the ocean for livelihood, subsitance, and way of life. Yet, many of our activities continiously pose a threat to marine populations and ecosystems. The implementation of Marine Protected Areas (MPAs), where extractive activities are regulated or prohibited on the basis of spatial deliniations, are one of many ways to manage the oceans. Current intenrational agreements call to achieve a certain amount of within-country conservation. Here we explore the benefits from implementing a market for conservation in which nations trade conservation towards a global target. We find that such a market reduces the costs of conservation by 46%, relative to a BAU scenario without a market where each country protects a fixed portion of their waters. When creating 12 distinct markets based on biogeography regionalization, we find that costs are reduced by 34.8%."
fontsize: 12pt
output:
  bookdown::pdf_document2:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: no
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
biblio-style: "nature"
header-includes:
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{pdfpages}
  - \usepackage{placeins}
---

@@

<!-- # Important sentences -->

<!-- Importantly, the quality of a patch can be defined independently of other pixels. This is true for habitat suitability, in which the abiotic characteristics that determine the suitability of a patch are independent of the protected status of any patch. -->

<!-- It is important that we acknowledge the equity concerns of any conservation intenrvention. Under the current approach to conservation, nations face two options: to conserve 30% in situ, or not engage in conservation at all. This market-based approach allows for a third option: it allows nations to decide how much to conserve in their waters. -->

<!-- Ocean grabbing, the term that relates to the loss of soveraignty in the ocean (see Benett) may be perceived as a problem. However, nations are not forced to conserve in situ, and will only do so if they are approrpiately compensated for it. -->

<!-- All ecosystems are euqlly important for the subsistance of nature and species that reside in them. But a coral reef is not comparable to a kelp forest.  A market design that doesn't acknowledge these differences may unintendendly bias conservation towards different ecosystems or geographic regions. When outcomes of a market are restricted to a geographic extent, such as air pollution markets, a buble policy is implemented. We borrow from this principle and use biogeographic regionalization principles to create ecologically-coherent market segments where nations trade similar habitat. For example, in the temperate eastern pacific, Mexico and the US will trade a same type of habitat, while in the Tropical Atlantic they wil. -->

<!-- Having tackled these three limitations (spatially-independent measure, x, and market bubbles), we proceed to simulate the trade outcomes of such a market. We evaluate the outcomes across a range of target coverages (1 - 100), and find that a market always results in reduced costs. The absolute and relative magnitude of the costs-savings changes with protection target, peaking at between X and Y, depending on of the choice market segmentation. Intuitively, increasing teh number of market segments reduces the opportunity for bargaining and reduces the gains from trade, but ensures the persistance of different habitats. -->

<!-- We pay especial attention to goals of interest, such as protecting 30% of the marine area (habitat suitability-weighted) within all national jurisdictions. We find that a market-based approach to conservation would reduce costs by X%, and that Y% less area would be needed. -->

# Main

More than three decades of conservation science have shown us that well-designed Marine Protected Areas (MPAs) can deliver large economic and ecological benefits\cite{halpern_2008,cabral_2019, cabral_2020}. Various international goals seek to protect anywhere between 10\% and 30\% of the global ocean with MPAs.  Along with these commitments, the scientific roadmap on where to site them\cite{roberts_2002,worm_2003,lucifora_2011,wallace_2011,albouy_2017,sala_2021} should be enough to see these goals realized. Why, then, have some nations achieved (and even surpassed) the ambitious 30\% target, while others lag behind even the modest 10\% target? Partly, because under the current unilateral conservation framework some nations face higher costs of conservation than others. Here, we show how an environmental market can be designed to lower the costs of conservation and induce cooperation towards a more efficient and equitable protection of the marine environment.

Habitat suitability, and the costs of protecting it, are heterogeneously distributed across the globe and non-perfectly correlated \cite{halpern_2008,roberts_2002,kroodsma_2018}. As a result, different nations can attain the same amount of conservation for dramatically different costs, questioning the efficiency and equity of the current unilateral approach to marine conservation. But what if there were a way to maximize global conservation while lowering national costs of conserving? We propose a new institution where nations facing large costs of conserving in their Exclusive Economic Zones (EEZ) can purchase conservation credits from nations with a surplus of habitat worth protecting, and show that the costs of conservation can be dramatically reduced.

The latest findings on marine conservation prioritization  show that a globally coordinated effort is more efficient than unilateral conservation efforts\cite{sala_2021}. While these results can point out where protection is most warranted and highlight the importance of coordination, they do little to incentivize nations to engage in it. Instead, institutions are needed to align actions with the desired conservation outcomes. One way to induce this desired cooperation is by allowing nations to conserve *in situ* (\emph{i.e.} within their EEZ) or pay to conserve *ex situ* (\emph{i.e.} in another nation's EEZ) to meet an aggregate target.

A market for conservation credits would allow just that. However, in designing such a market, two main challenges arise. First, not all ecosystems are created equally. Protecting one hectare of coral reef is not the same as protecting one hectare of kelp forest. A conservation market that ignores this could inadvertently fail to protect important habitat. Secondly, we live in a highly dynamic and connected world, one in which the biodiversity and costs of a patch of ocean depend on what else has been (or will be) protected, as well as it's own protection status. We provide solutions to both of these problems, and then simulate the trade outcomes of a global market for marine conservation credits and compare them to the business-as-usual baseline of unilateral conservation. For a range MPA coverage targets (1\% - 99\%), we find that a market always reduces the cost of conservation, resulting in a more efficient and equitable outcome.

# Designing a market for conservation credits

The first challenge in designing a market for conservation credits is to ensure that all ecosystems are properly protected. A coral reef can not be substituted for a kelp forest - they complement each other to produce our world's rich biodiversity. One way to avoid trading credits across different ecosystems is by using a "bubble policy" \cite{fr_1979}. In air quality management, this approach enables polluting plants to flexibly manage their air emissions. By placing a “bubble” over a city with multiple polluting plants and applying a single emission cap to the “bubbled” city, each plant can decide how to internally manage production and trade emissions permits accordingly.

We use biogeographic regionalization (Realms and Provinces \cite{spalding_2007} Fig S1) to create ecologically-coherent "blue bubbles" where nations trade conservation credits for similar habitats. For example, for Realm-based bubbles, Mexico and the United States trade conservation credits under two blue bubbles (Temperate Eastern Pacific [TEP] and Tropcial Atlantic [TA]), using bubble-specific conservation credits (\emph{e.g.} Kelp forest in the TEP and coral reefs in the TA).

The second challenge a market faces is in the deffinition of conservation targets. In the air pollution example, the target can be a fixed amount or ambient concentration of a pollutant, both of which are measurable quantities. Moreover, the amount of pollution in one bubble is not affected by the management status of the adjacent bubbles. However, biodiversity is an elusive variable, with multiple nuanced indices implicitly assigning weighs to different features (\emph{e.g.} species richness does not weigh abundance, while biomass does not acount for richness). Moreover, the biodiversity for any pixel in the world will be influenced by the protection status of itself, but also the network of connected pixels.

Instead, we define the value of a pixel based on its surface area and the suitability of the habitat it contains for the species distributed there. Habitat suitability is defined as X. Conveniently, the habitat suitability of any pixel in the ocean can be defined independently of the habitat suitability and protection status of other pixels. Habitat suitability maps for 4,000 (exact number here) marine species, as indicated by the probability of occurrence which is a function of X, Y and Z \cite{ready_2010}. The habitat suitability of any pixel in the world can be calculated as the average suitability across all species, conditional on the pixel being suitable (i.e. p > 0.5).

Marine Protected Areas are spatial conservation interventions that seek to directly protect habitat and, indirectly, the species that reside there. The siting of MPAs can therefore be guided by abiotic characteristics that are conducive to the persistence of species. Together, these abiotic characteristics define the habitat suitability of a patch for a given species.

# Simulations

We combine habitat suitability and fisheries revenue to generate supply curves for each nation, as well as a global biodiversity supply curve. While in practice this approach can be used for any given biodiversity target, I produce an example using the 30% area-based target (*i.e.* We translate this area-based target into biodiversity units). We then find the trading price in the global supply curve and use the country-level supply curves to calculate in-situ and ex-situ conservation for every country. We can then calculate the differences in costs, and estimate the gains from trade.

@@

The results presented so far assume that a unit of biodiversity is worth the same anywhere around the world. However, we might want to ensure that no single type of habitat or region is left out. To overcome this, markets can be established at different spatial scales. Here, we use different scales of zoning in the marine environment (*i.e.* Biogeographic zoning). \cite{spalding_2007} divided the oceans using a hierarchical approach with Realms (n = 12) containing Provinces (n = 60), which in turn contain Ecoregions (n = 219).

@@

![\label{fig:eezs-per-realm} Number of nations contained within each of the realms described by \cite{spalding_2007}. The vertical dashed line lies at 2,  the minimum number of nations required for trade to occur.](results/img/eezs_per_realm.png)

@@

![\label{fig:realm-map}Map of the 12 realms, as defined by \cite{spalding_2007}.](results/img/realm_map.png)

Map of provinces

@@

We must first identify the costs and benefits of conservation. Conserving biodiversity may provide benefits to society. For example, areas with higher species richness and biomass are more attractive to tourism \cite{williams_2000}. Species richness also leads to increased stability of ecosystems \cite{tilman_1995}, and may reduce exposure to pathogens \cite{lamb_2017}. The quantification is not straightforward, but we can assume that biodiversity indices are positively correlated with their benefits they provide (although saturation may occur past a given point).

@@

On the other hand, we need to quantify the costs associated to each pixel. An often-cited cost of implementing MPAs is that of losses in fisheries revenues \cite{smith_2010}. In this case, we define costs as losses in fisheries landings ^[I intend to use losses in fisheries revenues in the future. This will involve species-level changes in landings, as well as species-level price data. Essentially, the global losses in landings used here will be weighed by species price data.]. Since protection of some areas may lead to rebuilding fisheries, parts of the ocean might have negative costs. This means that there are gains to be made in conserving them. A map of global costs of conservation is shown in Figure \ref{fig:landing-losses}. As before, these data do not consider the transient dynamics, and represent the equilibrium state of protecting any given pixel.

@@

Although there are important caveats, these data provide us with a way of calculating the costs and benefits of protecting any place under a nation's jurisdiction. While protecting the high seas may provide many benefits \cite{white_2014, sumaila_2015}, our market-based approach requires enforceable property rights, such as those offered by Exclusive Economic Zones. Moreover, biodiversity benefit is higher within Exclusive Economic Zones, where most of the species live \cite{sala_2021}. Therefore, we will ignore areas beyond national jurisdiction.

@@

<!-- MAYBE:: A meta-analysis of simulation studies suggests that targets ranging from biodiversity protection to food provisioning can be indeed achieved by protecting ~30% of our coastal oceans \cite{oleary_2016,cabral_2019}. However, these area-based targets ultimately imply a desired level of biodiversity benefit. -->
